\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{url}
\usepackage{tikz}
\usetikzlibrary{positioning,arrows.meta}
\usepackage{pgfplots}
\pgfplotsset{compat=1.16}
\usepackage[hidelinks]{hyperref}

\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\um}{\ensuremath{\times}}

\begin{document}

\title{JacInteropBench: Characterizing the Marshalled Boundaries of a
Multi-Target Compiler with a Differential-Identity Oracle}

\author{\IEEEauthorblockN{Anonymous Author(s)}
\IEEEauthorblockA{\textit{Jaseci Labs} \\
\texttt{interopbench@jaseci.org}}}

\maketitle

\begin{abstract}
Modern applications are polyglot by construction: logic in Python,
presentation in JavaScript, and hot paths in native or WebAssembly code, stitched
together across process, language, and serialization boundaries that no single
compiler can see through. Jac is a \emph{synechic} language that compiles one
Python-like syntax to Python bytecode, JavaScript, and native machine code, and
therefore owns both sides of every cross-tier call. Owning both sides makes the
cost of a boundary a first-class, measurable property of the language rather than
an emergent property of hand-written glue. We present \textbf{JacInteropBench}, a
microbenchmark suite that characterizes the \emph{marshalled} boundaries of this
compiler---where a call becomes an RPC, an FFI thunk, or a WebAssembly
import/export and every crossing value is serialized. The suite organizes ten
measured kernels across two harness families (in-process native bridges and
cross-runtime hosts), anchored by two named free floors, and pairs every
marshalled cell with an exact reference twin that runs the same deterministic
operation without the crossing. Its load-bearing contribution is not a headline
speedup but a \emph{differential-identity oracle}: byte-identical digests across
each twin group are a correctness and compiler-regression witness, gated in
continuous integration, while timing is reported strictly as ``this microprogram,
at this size, on this machine.'' On a 16-core x86\_64 host, an
eight-point work sweep separates the two coefficients that a single size
conflates: execution placement is a $\approx$$42\times$ per-iteration
\emph{slope} ($4.4$ vs.\ $184$\,ns/iter, \code{na} vs.\ \code{sv}), while the
\code{native$\rightarrow$server} callback crossing is a fixed
$\approx$$5.4$\,\textmu s per-call \emph{intercept}---so its headline
$1.17\times$ ratio is a constant seen at one small size and decays to
$1.01\times$ once the callee does real work. A complementary payload sweep over the same loopback RPC, from $1$ to $100{,}000$
returned elements, resolves the crossing into a $\approx$$19$\,ms fixed cost plus
a $934$\,ns/element serialization slope (95\% bootstrap CI $912$--$999$)---$3.1\times$ the in-process list's
$298$\,ns/element---with byte-identical digests at every size, and directly
observes that slope overtake the fixed cost at $N\approx21{,}000$. We further report
low-nanosecond \code{C}-ABI foreign-call costs and, at median-of-seven with
reported spread, cross-runtime floors of $10$--$213\times$ the direct dispatch
they replace. The
oracle is not Jac-specific: twin-anchored differential identity is
adoptable by any toolchain that generates both sides of a boundary---gRPC
codegen, the WebAssembly component model, or a polyglot VM---and here receives a
complete instantiation.
\end{abstract}

\begin{IEEEkeywords}
Workload characterization, language interoperability, foreign function interface,
WebAssembly, RPC, microbenchmarking, differential testing
\end{IEEEkeywords}

\section{Introduction}

Open any well-built product and count the languages a maintainer must read:
Python and SQL where the data lives, TypeScript and JSX at the edge, C or Rust in
the hot path, and a layer of JSON, protobuf, and OpenAPI documents wiring them
together. Every architectural seam is also a change of language, type system, and
serialization format, and \emph{no compiler can see across any of them}. The
whole-program type checker of the modern stack is \code{grep}.

Jac~\cite{jac} is a bet against that arrangement. It compiles one clean,
Python-like syntax to Python bytecode (its \emph{server}, or \code{sv},
codespace), to JavaScript (its \emph{client}, or \code{cl}, codespace), and to
native machine code and WebAssembly (its \emph{native}, or \code{na}, codespace),
with the entire PyPI, npm, and C ecosystems reachable through a plain
\code{import}. We call this property \emph{synechic} (from the Greek
\emph{synecheia}, continuity): one continuous, statically checked medium across
ecosystems and tiers, so that renaming a field breaks every stale use---server,
client, or native---at compile time rather than in production.

Continuity does not delete the boundaries that are physics. A cross-tier call is
still an RPC because the network is real; a native call still crosses an FFI ABI;
a browser kernel still crosses a WebAssembly import table. What synechic
compilation changes is \emph{who owns the crossing}: the compiler emits both
sides, so the cost of each boundary becomes a measurable, attributable property
of the language rather than an emergent property of hand-written glue. That
measurability is an opportunity and an obligation. If Jac claims that its
\code{\_to\_wire}/\code{\_\_from\_wire} marshalling is cheap, that its
\code{NativeListView} is zero-copy, that its \code{C} struct-by-value ABI is
correct, and that its WebAssembly exports are fast, then each claim needs a
number behind it and a regression guard around it. Before this work, none did.

Jac already ships \code{ownbench}, a microbenchmark suite for the \emph{free}
\code{na$\rightarrow$na} boundary and the memory-management dial (ownership,
regions, GC). Everything \code{ownbench} measures stays inside one runtime. It
never exercises a marshalled crossing. \textbf{JacInteropBench} is the
complementary suite: one benchmark scenario per boundary kernel, an
\code{ns=}-plus-digest output contract, a Jac-native harness, a
differential-identity oracle, and continuous-integration regression gating,
extended from \code{ownbench}'s single-runtime discipline to the six marshalled
cells of Jac's interop matrix.

This makes Jac the \emph{experimental system}, not the result. The question we
characterize is: \emph{what are the performance characteristics of application
workloads that cross browser, managed-runtime, and native-code boundaries, and
what does owning both sides of every crossing let a compiler measure that
hand-written glue hides?} We treat each boundary as a distinct workload class
with a distinct dominant cost (Table~\ref{tab:profiles}), and we hold the
\emph{language} fixed while varying the \emph{boundary}---the inverse of a
cross-language benchmark game, which holds the problem fixed and varies the
language.

We make three contributions:

\begin{enumerate}
  \item \textbf{A boundary taxonomy and kernel map} (Section~\ref{sec:matrix})
  that decomposes Jac's cross-codespace surface into ten measured kernels across
  two harness families plus two named free floors, and pairs each boundary with
  its dominant cost and the compiler strategy that cost enables.
  \item \textbf{A twin-anchored differential-identity methodology}
  (Section~\ref{sec:method}) in which every marshalled cell names an exact
  reference twin, byte-identical canonical digests witness semantic equivalence
  and compiler regressions, and timing is reported only within an explicitly
  bounded scope of valid conclusions.
  \item \textbf{A characterization} (Section~\ref{sec:results}) of native-bridge
  and cross-runtime boundary costs on a commodity x86\_64 host, including an
  eight-point work sweep and a sixteen-point payload sweep that decompose per-call
  time into a crossing intercept and an execution (or serialization) slope,
  together with a frank threats-to-validity account
  (Section~\ref{sec:threats}) that states exactly what each number does and does
  not license.
\end{enumerate}

\section{Background: The Interop Matrix}
\label{sec:matrix}

Jac programs are partitioned into three \emph{codespaces}---server (\code{sv},
Python bytecode), client (\code{cl}, JavaScript), and native (\code{na}, machine
code / WebAssembly)---plus the ambient \code{C} and Python (\code{py})
ecosystems reachable by import. A call within one codespace is \emph{free}: no
value is serialized. A call \emph{between} codespaces, or out to \code{C} or the
browser, is \emph{marshalled}: it becomes an RPC, an FFI thunk, or a WebAssembly
import/export, and every value that crosses is serialized and reconstructed on
the far side. Figure~\ref{fig:overview} shows the surface: which edges the
compiler emits for free, and which it must marshal---and thus must be measured.

\begin{figure}[t]
\centering
\begin{tikzpicture}[
  font=\small,
  space/.style={draw, rounded corners, minimum width=15mm, minimum height=8mm,
                align=center, thick},
  ambient/.style={draw, dashed, rounded corners, minimum width=13mm,
                  minimum height=7mm, align=center},
  free/.style={-{Latex}, thick},
  marsh/.style={-{Latex}, thick, dotted},
]
\node[space] (sv) at (0,0) {\code{sv}\\\scriptsize Python};
\node[space] (cl) at (-2.6,-2.2) {\code{cl}\\\scriptsize JS};
\node[space] (na) at (2.6,-2.2) {\code{na}\\\scriptsize native/wasm};
\node[ambient] (py) at (-2.6,1.4) {\code{py}\\\scriptsize PyPI};
\node[ambient] (c)  at (2.6,1.4)  {\code{C}\\\scriptsize libs};

% free edges (solid)
\draw[free, {Latex}-{Latex}] (sv) -- (py) node[midway,left=-1pt] {\scriptsize free};
\draw[free, {Latex}-{Latex}] (na) to[out=-50,in=-90,looseness=8]
  node[below=-1pt] {\scriptsize free} (na);

% marshalled edges (dotted)
\draw[marsh, {Latex}-{Latex}] (sv) -- (na)
  node[midway,right=1pt] {\scriptsize JIT / \code{na\{\}} / FFI};
\draw[marsh, {Latex}-{Latex}] (cl) -- (sv)
  node[midway,left=1pt] {\scriptsize HTTP+JSON};
\draw[marsh, {Latex}-{Latex}] (cl) -- (na)
  node[midway,below] {\scriptsize wasm};
\draw[marsh, {Latex}-{Latex}] (na) -- (c)
  node[midway,right=-1pt] {\scriptsize FFI ABI};
\draw[marsh, {Latex}-{Latex}] (sv) to[out=60,in=120,looseness=8]
  node[above] {\scriptsize RPC ($\mu$svc)} (sv);
\end{tikzpicture}
\caption{Jac's codespaces and ambient ecosystems. Solid arrows are
\emph{free} in-codespace calls (nothing serialized); dotted arrows are the
\emph{marshalled} boundaries JacInteropBench characterizes---each becomes an
RPC, an FFI thunk, or a WebAssembly import/export.}
\label{fig:overview}
\end{figure}

\begin{table}[t]
\caption{Boundary $\rightarrow$ kernel map. Prefix \code{iop\_*} = in-process
native bridge (family~1), \code{xop\_*} = cross-runtime (family~2),
\code{base\_*} = a named free floor. The ``real-world stand-in'' column names
each kernel's design target and analogue; as implemented the enabled cells are
synthetic microprograms, so that column is \emph{aspirational}---though the
callee-work sweep (Section~\ref{sec:sweep}) and the \code{xop\_feed\_payload}
payload sweep (Section~\ref{sec:payload}) now realize the two scaling axes.}
\label{tab:map}
\centering
\small
\begin{tabular}{@{}llll@{}}
\toprule
Boundary & Wire & Kernel(s) & Real-world stand-in \\
\midrule
\code{sv}$\rightarrow$\code{sv} & free & \code{base\_call} & Python-call floor \\
\code{sv}$\leftrightarrow$\code{py} & free & \code{base\_pyimport} & NumPy/sklearn floor \\
\code{sv}$\rightarrow$\code{na} & JIT & \code{iop\_call} & Python calls native \\
\code{na}$\rightarrow$\code{sv} & \code{na\{\}} & \code{iop\_cb} & native calls back \\
\code{sv}$\leftrightarrow$\code{na} & JIT & \code{iop\_symmetric} & placement dial \\
\code{na}$\leftrightarrow$\code{C} & FFI ABI & \code{iop\_ffi\_scalar} & libm \\
& & \code{iop\_ffi\_struct} & raylib, SQLite \\
& & \code{iop\_ffi\_vtable} & CEF callbacks \\
\code{sv}$\rightarrow$\code{sv} & HTTP+JSON & \code{xop\_svc\_split} & billing fan-out \\
& HTTP+JSON \code{[N]} & \code{xop\_feed\_payload} & typed feed payload \\
\code{cl}$\leftrightarrow$\code{sv} & HTTP+JSON & \code{xop\_feed} & littleX feed \\
\code{cl}$\rightarrow$\code{na} & wasm export & \code{xop\_wasm\_call} & PolyBench in browser \\
\bottomrule
\end{tabular}
\end{table}

Table~\ref{tab:map} lists the surface. The split into two harness families is not
cosmetic: it follows from what a boundary requires of the measurement harness. A
native bridge lives \emph{inside} one process---an in-process JIT or an
ahead-of-time native binary---so a command adapter can drive it directly. A
cross-runtime boundary requires a \emph{second} runtime---a live Jac server, a
generated JavaScript client, a Node WebAssembly host---so a host adapter must own
that runtime's lifecycle: allocate a port, start it, poll readiness, capture
logs, reset state between variants, and terminate-then-kill in every exit path.
The two families share one parsing, aggregation, and oracle module
(\code{common.jac}) but keep their build and host lifecycles separate.

Table~\ref{tab:profiles} is the conceptual spine of the characterization: it
names, per boundary, the \emph{dominant cost} a workload pays and the compiler
strategy that cost licenses. It is a taxonomy, not a measured policy---we do not
yet implement automatic strategy selection---but it states what a synechic
compiler is positioned to decide once each cost is measured, and it organizes
which cost each kernel in Table~\ref{tab:map} is designed to expose.

\begin{table}[t]
\caption{Boundary cost profiles. The dominant cost defines the workload class;
the enabled strategy is what owning both sides of the crossing lets a compiler
choose. Strategy selection is design-implication (Section~\ref{sec:implications}),
not a measured result.}
\label{tab:profiles}
\centering
\small
\begin{tabular}{@{}lll@{}}
\toprule
Boundary & Dominant cost & Enabled strategy \\
\midrule
\code{cl}$\rightarrow$\code{sv} & network + JSON ser. & batch / migrate / cache \\
\code{sv}$\leftrightarrow$\code{na} & FFI transition + marshal & inline vs.\ copied thunk \\
\code{na}$\rightarrow$\code{sv} & callback + runtime entry & hoist / amortize round trip \\
\code{sv}$\rightarrow$\code{sv} & RPC + placement & co-locate vs.\ split \\
\code{na}$\leftrightarrow$\code{C} & ABI class (register/\code{byval}) & pass-by-value vs.\ pointer \\
\code{cl}$\rightarrow$\code{na} & wasm import + BigInt copy & keep compute in-runtime \\
\bottomrule
\end{tabular}
\end{table}

\section{Methodology}
\label{sec:method}

\subsection{The kernel contract}

Every kernel obeys a load-bearing output contract inherited from \code{ownbench}
and extended for boundary crossings. Each invocation prints exactly one non-empty
\emph{canonical digest} line, exactly one effective \code{ns=$\langle$integer$\rangle$}
timing line, and zero or more unique integer \code{m:$\langle$key$\rangle$=$\langle$value$\rangle$}
metric lines. The parser strips both \code{ns=} and \code{m:} lines before
comparing digests, and rejects missing or multiple timing lines, duplicate
metrics, a changing metric-key set, or a nondeterministic digest.

Determinism is enforced by construction: fixed linear-congruential-generator
(LCG) seeds, sorted iteration, fresh or reset state per variant, and no clocks,
generated identifiers, transport envelopes, or object addresses in the digest.
Each run selects exactly one named work size, so a single run reports per-call or
per-batch time \emph{at that size only}. Invocations within a run are interleaved
per variant (paired sampling) so that machine drift hits both sides of a
comparison evenly.

\subsection{The differential-identity oracle}

The suite's central discipline is that \emph{identity is never checked against a
generic floor}. Every marshalled cell names an exact reference twin that runs the
same operation over the same deterministic input without the crossing. The
\code{free} variant of \code{iop\_call} runs an LCG checksum in ordinary
\code{sv}; the \code{bridge} variant runs the identical checksum inside a
\code{na\{\}} block. Both must emit the byte-identical canonical digest
\code{call:263997602}. When they do, three facts are witnessed simultaneously:
the marshalling preserved semantics, the native code path is functionally
equivalent to the server path, and no compiler change silently perturbed either.
This is the load-bearing contribution. A speedup is interesting; a
byte-identical digest across a marshalled seam, checked on every commit, is a
guarantee.

The oracle's reach is best seen in one number. The same deterministic LCG
checksum, \code{263997602}, is reproduced byte-for-byte across six kernels
spanning every wire in the suite---\code{sv} and \code{na} execution
(\code{iop\_call}, \code{iop\_symmetric}), a native$\rightarrow$server callback
(\code{iop\_cb}), two HTTP+JSON RPC paths (\code{xop\_feed},
\code{xop\_svc\_split}), and a \code{wasm32} export (\code{xop\_wasm\_call}).
One workload, six runtimes and transports, one canonical value: that identity is
the correctness claim the whole suite is built to defend.

Concretely, the continuous-integration gate (\code{ci\_bridges.sh}) runs the
enabled kernels at small sizes, filters out both \code{ns=} and \code{m:} lines,
and asserts canonical-digest identity across each declared twin group plus a set
of named structural invariants (Section~\ref{sec:audit}). \emph{Timing values
never gate CI.} A regression is a semantic divergence or a broken structural
invariant, never a slow number on a noisy runner. The same scalar pair also
lives as a standalone compiler regression test
(\code{test\_interopbench\_bridges.jac}) that invokes both paths through
\code{python -m jaclang run} and asserts digest identity, with no GNU-time,
NumPy, C-compiler, Node, server, or WebAssembly prerequisite.

\subsection{Structural audits}
\label{sec:audit}

Beyond digest identity, the suite records named \emph{structural} facts about the
generated code (\code{interop\_audit.json}). Manifest/wrapper audits are enabled
for the two mixed-JIT scalar cells (\code{iop\_call}, \code{iop\_cb}); the FFI
cells' IR/ABI assertions live in the compiler's existing owner tests
(\code{test\_abi\_lib}, \code{test\_native\_gen\_pass}) rather than in this
suite. For \code{iop\_call}, the audit
asserts the \code{InteropManifest} exports exactly \code{native\_checksum} and no
imports, that the generated Python wrapper contains exactly one
\code{get\_function\_address('native\_checksum')} lookup and one matching
\code{ctypes.CFUNCTYPE} signature, and that the pure-\code{sv} baseline
(\code{sv\_checksum}) generates \emph{no} native binding. For \code{iop\_cb}, it
additionally asserts one native import (\code{sv\_callback\_checksum}) and its
\code{\_\_jac\_interop\_py\_funcs\_\_} registration. These are compiler-drift
sentinels: they fail loudly if a codegen change adds a stray ctypes stub or
drops a trampoline, catching regressions that a purely behavioral oracle would
miss. Table~\ref{tab:audit} shows the concrete asserted facts; all matched on the
measured build.

\begin{table}[t]
\caption{Structural-audit facts (\code{interop\_audit.json}), expected =
actual on the measured build. Enabled for the two mixed-JIT scalar cells.}
\label{tab:audit}
\centering
\footnotesize
\setlength{\tabcolsep}{4pt}
\begin{tabular}{@{}lll@{}}
\toprule
Cell & Asserted fact & Count \\
\midrule
\code{iop\_call} & \code{native\_exports}=\{\code{native\_checksum}\} & 1 \\
                 & \code{get\_function\_address} lookup & 1 \\
                 & \code{ctypes.CFUNCTYPE} signature & 1 \\
                 & baseline \code{sv\_checksum} native bindings & 0 \\
\midrule
\code{iop\_cb}   & \code{native\_imports}=\{\code{sv\_callback\_checksum}\} & 1 \\
                 & \code{\_\_jac\_interop\_py\_funcs\_\_} reg. & 1 \\
                 & baseline \code{sv\_direct\_checksum} bindings & 0 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Provenance and spread}

Results are schema version~2. Every document carries a \code{provenance} block
(host, platform, machine, CPU count, Python version, UTC capture time) so a
number is never quoted without the machine it came from. Each cell records
\code{samples}, \code{median\_ns}, \code{min\_ns}, \code{max\_ns},
\code{stdev\_ns}, and \code{iqr\_ns}. What is deliberately \emph{not} computed:
confidence intervals, ratio uncertainty, or an outlier policy---spread is
reported, not gated. Results also label \code{command\_adapter},
\code{timing\_source}, and \code{rss\_scope}; cross-runtime cells that cannot
share a clock label \code{timing\_source} \emph{per variant} rather than stamping
one label over a mixed pair.

\section{Kernels}
\label{sec:kernels}

\paragraph{Family~1: native bridges (\code{iop\_*})}
\code{iop\_call} is the vertical slice: one mixed source runs a deterministic
scalar LCG checksum in \code{sv} (\code{free}) and in \code{na\{\}}
(\code{bridge}). It is an \emph{execution-placement} benchmark---the ratio blends
the \code{sv$\rightarrow$na} crossing with native-vs-server execution speed, not
isolated boundary cost. \code{iop\_cb} isolates the crossing: both variants run
the work in \code{sv}, and only the \code{bridge} path adds a
native$\rightarrow$server callback round trip, making it the cleanest boundary
cell. \code{iop\_symmetric} sweeps all four placements (\code{sv\_local},
\code{sv\_to\_na}, \code{na\_local}, \code{na\_to\_sv}) over identical callee
work. The three \code{iop\_ffi\_*} kernels exercise the \code{na$\leftrightarrow$C}
ABI through a deterministic \code{C} fixture: \code{iop\_ffi\_scalar} churns
\code{libm} \code{sqrt}; \code{iop\_ffi\_struct} passes
\code{\_Static\_assert}ed 4/12/16/24/44-byte structs by value across System~V
INTEGER and MEMORY classes; \code{iop\_ffi\_vtable} has \code{C} invoke a Jac
callback through a \code{\_\_clibcb} trampoline.

\paragraph{Family~2: cross-runtime (\code{xop\_*})}
\code{xop\_svc\_split} places an \code{sv import from billing} between two
loopback \code{jac start} deployments, with direct in-process dispatch as its
twin. \code{xop\_feed\_payload} uses the same loopback \code{sv$\rightarrow$sv}
seam but varies the returned \code{list[$N$]} cardinality rather than callee work,
isolating a per-element serialization slope (Section~\ref{sec:payload}).
\code{xop\_feed} drives a scalar provider through a generated JavaScript
client over HTTP+JSON. \code{xop\_wasm\_call} compiles an integer kernel to
\code{wasm32} and invokes it from a Node host, against a native bridge as
reference. As Section~\ref{sec:threats} details, the scalar cross-runtime cells
are honest \emph{floors}, not microservice or same-source native-vs-wasm
measurements; \code{xop\_feed\_payload} adds the one payload-scaling curve.

\section{Results}
\label{sec:results}

All numbers below were captured on host \code{zaitop}: Linux~6.18 x86\_64,
16~cores, CPython~3.14.6. Single-size tables use the small work size
(\code{args = 200, 20}); Section~\ref{sec:sweep} sweeps eight callee-work sizes
and Section~\ref{sec:payload} sixteen payload cardinalities. Family-1 reports the
median of five interleaved invocations, family-2 the median of seven (five to nine for
the payload sweep). Every cell's variants produced byte-identical canonical
digests---the oracle passed for all ten measured kernels at every size.

\subsection{Native bridges}

Table~\ref{tab:bridges} reports the family-1 results, anchored by the named
free floor \code{base\_call}: a bare \code{sv$\rightarrow$sv} call over the same
LCG leaf measures $34{,}853$\,ns/call at this size and emits the same
\code{call:263997602} digest as the rest of the suite. So \code{iop\_call}'s
$36{,}365$\,ns \code{free} reference is essentially the bare \code{sv}-call floor
plus noise---the \code{sv$\rightarrow$na} comparisons below are against a real
floor, not inflated server code. (The companion \code{sv$\leftrightarrow$py}
floor \code{base\_pyimport} is defined but requires NumPy in the \code{jac}
runtime environment, which this host does not provision; we report it as
env-gated rather than fabricate a number.) Three findings stand out.

\emph{Native execution is fast, but that is placement, not boundary cost.}
\code{iop\_call} runs the LCG $8.5\times$ faster inside \code{na\{\}}
($4{,}282$\,ns/call) than in \code{sv} ($36{,}365$\,ns/call), and
\code{iop\_symmetric}'s \code{na\_local} floor ($1{,}140$\,ns/call) is $31\times$
below its \code{sv\_local} twin. These ratios are dominated by how fast each
runtime executes the kernel, not by what the crossing costs.

\emph{The crossing itself is cheap.} \code{iop\_cb}, which holds execution in
\code{sv} and adds only a native$\rightarrow$server callback round trip, costs
$40{,}466$\,ns/call against a $34{,}513$\,ns/call reference---a $1.17\times$
ratio, i.e.\ the round trip adds roughly $17\%$. Over the five samples the two
distributions do not overlap (reference \code{max} $721{,}184 <$ bridge
\code{min} $754{,}811$\,ns), and the min/max-bracketed ratio runs
$1.05$--$1.23\times$; the effect is real, and we quote it with that range rather
than as a bare point. This is the suite's cleanest statement of pure boundary
overhead.

\emph{The C ABI is close to free for scalars and callbacks---but not zero.} The
integer per-call metric (\code{m:per\_ffi\_call\_ns}) floors both FFI scalar and
trampoline costs to $2$--$3$\,ns, which understates a real sub-nanosecond
premium visible in the batch medians: \code{iop\_ffi\_vtable}'s
\code{C$\rightarrow$Jac} trampoline adds $11.9\%$ to batch wall time
($11{,}599$ vs.\ $10{,}363$\,ns over the same call count), and
\code{iop\_ffi\_scalar}'s \code{sqrt} adds $21.5\%$ over its arithmetic reference
($12{,}243$ vs.\ $10{,}077$). The trampoline is cheap, not free; we read it off
the batch wall time, not the quantized per-call integer. Struct-by-value is
where marshalling shows plainly: \code{iop\_ffi\_struct} makes ten \code{C} calls
per outer call across five heterogeneous shapes, averaging $15$\,ns per foreign
call---an average across INTEGER- and MEMORY-class shapes, not a per-shape cost,
and the reason its batch median ($616{,}582$\,ns) sits $6.8\times$ above the
computed reference.

\emph{Run-to-run variance is placement-dependent, not crossing-dependent.} The
sharpest workload observation in the family-1 data is in the dispersion, not the
medians. Every \code{sv}-resident variant is noisy---\code{iop\_call} free
\code{stdev} $24{,}307$\,ns, \code{na\_to\_sv} $46{,}539$, \code{iop\_cb} bridge
$37{,}081$---while every \code{na}-resident variant is nearly silent:
\code{iop\_call} bridge \code{stdev} $894$, \code{na\_local} $374$, and the
\code{iop\_ffi\_*} reference cells $15$--$196$\,ns. Run-to-run jitter tracks
\emph{where the work executes} across roughly $1.5$ orders of magnitude: CPython
scheduling, not the marshalled seam, dominates variance. This is why paired,
interleaved sampling matters---it puts that CPython jitter on both sides of every
comparison evenly. We report \code{min}/\code{max}/\code{stdev}/\code{iqr} per
cell and gate on none of them (Section~\ref{sec:method}); the native-path ratios
above are read off medians whose spread ($<3\%$ IQR) is small enough not to
swallow them.

\begin{table}[t]
\caption{Family~1 native bridges. Median wall time and the kernel's own
per-call metric (\code{ns}). Digests identical within every cell.}
\label{tab:bridges}
\centering
\small
\setlength{\tabcolsep}{4.5pt}
\begin{tabular}{@{}llrr@{}}
\toprule
Cell & Variant & Median (ns) & per-call (ns) \\
\midrule
\code{iop\_call} & free (\code{sv} ref) & 727{,}310 & 36{,}365 \\
                 & bridge (\code{na})   & 85{,}648  & 4{,}282 \\
\midrule
\code{iop\_cb}   & free (\code{sv} ref)        & 690{,}274 & 34{,}513 \\
                 & bridge (\code{sv}$\to$\code{na}$\to$\code{sv}) & 809{,}333 & 40{,}466 \\
\midrule
\code{iop\_symmetric} & sv\_local & 713{,}816 & 35{,}690 \\
                      & sv\_to\_na & 92{,}160 & 4{,}608 \\
                      & na\_local  & 22{,}818 & 1{,}140 \\
                      & na\_to\_sv & 778{,}104 & 38{,}905 \\
\midrule
\code{iop\_ffi\_scalar} & ref (int math)   & 10{,}077 & 2 \\
                        & ffi (\code{sqrt}) & 12{,}243 & 3 \\
\midrule
\code{iop\_ffi\_struct} & ref (computed)   & 90{,}482  & 2 \\
                        & ffi (struct ABI) & 616{,}582 & 15 \\
\midrule
\code{iop\_ffi\_vtable} & ref (direct Jac)      & 10{,}363 & 2 \\
                        & ffi (\code{C}$\to$Jac) & 11{,}599 & 2 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Cross-tool FFI baseline}
\label{sec:xtool}

A per-foreign-call cost is only interpretable against a comparand. The
\code{iop\_ffi\_scalar} kernel calls \code{libm}'s \code{sqrt} from Jac's native
codespace; Jac's mixed-JIT path emits a \code{ctypes.CFUNCTYPE} thunk
(Section~\ref{sec:audit}), so the natural controlled A/B is the \emph{same} libm
\code{sqrt}, over the \emph{same} deterministic loop, called from Python's own
two standard FFI mechanisms---\code{ctypes} and \code{cffi}---plus each side's
no-FFI arithmetic floor. Every variant reconstructs the byte-identical digest
\code{sqrt:402000}: the differential-identity oracle now spans \emph{toolchains},
not just codespaces, so ctypes, cffi, and Jac-native are provably computing one
value from the same foreign calls.

Table~\ref{tab:xtool} reports median per-foreign-call cost (medians over five
samples; per-call cost is essentially flat in $n$ because each \code{sqrt} is
$O(1)$). Jac's native FFI ($3.1$\,ns) is $\mathbf{102\times}$ cheaper than
\code{cffi} ($314$\,ns) and $\mathbf{153\times}$ cheaper than \code{ctypes}
($475$\,ns) for the identical foreign call---and, strikingly, $12\times$ cheaper
than \emph{pure Python arithmetic with no FFI at all} ($39$\,ns). This is the
paper's one cross-tool result, and it is deliberately scoped: it compares
Jac-native FFI against the Python-side FFI mechanisms Jac's own bridge emits, on
one libm scalar. It is \emph{not} a comparison against pybind11, PyO3, or a
hand-written C extension---those need matched kernels and per-system warm-up and
are the subject of the deferred verdict paper (Section~\ref{sec:related}).

\begin{table}[t]
\caption{Cross-tool FFI: same \code{libm} \code{sqrt}, same loop, same digest
\code{sqrt:402000}, median per foreign call (ns), $n{=}200$. Jac-native vs.\ the
Python FFI mechanisms its own bridge emits.}
\label{tab:xtool}
\centering
\small
\begin{tabular}{@{}llr@{}}
\toprule
Toolchain & Path & ns/call \\
\midrule
Jac \code{na} & FFI (\code{sqrt})      & $3.1$ \\
Jac \code{na} & ref (no FFI)           & $2.5$ \\
Python \code{cffi}   & FFI (\code{sqrt}) & $314$ \\
Python \code{ctypes} & FFI (\code{sqrt}) & $475$ \\
Python & ref (no FFI)                  & $39$ \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Cross-runtime}

Table~\ref{tab:xruntime} reports the family-2 results, now medians of
\emph{seven} interleaved invocations per variant with reported spread---enough to
replace the earlier single-warm-up floors with ranged estimates, though still
short of stable V8 tiering. The marshalled crossings are two-plus orders of
magnitude more expensive than the direct dispatch they replace, as physics
demands: a loopback HTTP+JSON RPC (\code{xop\_svc\_split}) costs $135\times$ its
in-process twin and a generated-client feed (\code{xop\_feed}) $213\times$ its
direct provider call, while a \code{wasm32} export (\code{xop\_wasm\_call}) is
$10\times$ a native bridge under an asymmetric host loop. The crossing-side IQRs
are wide ($11$--$14\%$; e.g.\ \code{xop\_feed} client $\pm18.7$\,M\,ns on a
$148$\,M\,ns median), so we still read these as order-of-magnitude \emph{floors},
not precise ratios---and they remain the least apples-to-apples comparisons in
the suite, since the two sides do not share a clock (each carries its own
\code{timing\_source}). Every cross-runtime cell's digests were byte-identical
across all seven samples.

\begin{table}[t]
\caption{Family~2 cross-runtime. Direct twin vs.\ marshalled crossing, median
of \emph{seven} interleaved invocations; \code{[min, max]} in brackets. Timing
sources differ per variant (see text); crossing-side spread is wide, so ratios
are order-of-magnitude floors.}
\label{tab:xruntime}
\centering
\small
\begin{tabular}{@{}lrrr@{}}
\toprule
Cell & Direct med.\ (ns) & Crossing med.\ (ns) & Ratio \\
\midrule
\code{xop\_feed}      & 696{,}877 & 148{,}257{,}648 & $213\times$ \\
\code{xop\_svc\_split} & 689{,}606 & 93{,}013{,}026  & $135\times$ \\
\code{xop\_wasm\_call} & 16{,}910  & 170{,}557       & $10\times$ \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Separating fixed cost from callee work}
\label{sec:sweep}

A single work size cannot tell a fixed per-call crossing cost apart from
per-iteration execution cost. To separate them we sweep the callee work
parameter across eight sizes ($w\in\{25,\dots,3200\}$ LCG iterations/call at
$40$ calls, five timed samples per point through the same \code{jac run} adapter)
and fit each variant's per-call time to
$t(w)\approx t_0 + s\cdot w$. The intercept $t_0$ is the per-call
boundary/dispatch cost independent of callee work; the slope $s$ is the
per-iteration execution cost. All fits are near-perfect lines
($R^2\ge0.999$ for every \code{sv}-resident series; Table~\ref{tab:fits}).

Two facts fall straight out. First, \emph{slope is bimodal by placement}:
\code{na}-resident execution runs at $\approx$$4.4$\,ns/iter
(\code{iop\_call} bridge, \code{na\_local}, \code{sv\_to\_na}) while
\code{sv}-resident execution runs at $\approx$$184$\,ns/iter---a
$\approx$$42\times$ gap that is a property of \emph{where the work runs}, not of
the crossing. Second, and more decisively, \emph{the callback boundary is a
constant, not a percentage}. In \code{iop\_cb} the \code{free} and \code{bridge}
slopes are statistically identical ($182.2$ vs.\ $183.5$\,ns/iter, within
$0.7\%$): the two variants do the same work, so their fitted lines are parallel
and separated by a fixed intercept gap of $\approx$$5.37$\,\textmu s---the pure
per-call callback cost. That constant is exactly why the headline
\code{iop\_cb} ratio decays as work grows (Figure~\ref{fig:decay}): from
$1.73\times$ at $w{=}25$ through $1.12\times$ at $w{=}200$ down to $1.01\times$
at $w{=}3200$. The much-quoted ``$1.17\times$'' is not a property of the
boundary; it is a fixed $5.4$\,\textmu s cost measured at one small size, and it
vanishes into the noise once the callee does real work. \code{iop\_symmetric}
makes the decomposition visible in one cell: \code{sv\_to\_na} runs at the
\emph{native} slope ($4.42$) yet carries a $\approx$$3.2$\,\textmu s crossing
intercept, so entering \code{na} pays a one-time per-call toll and then executes
$42\times$ faster per iteration---precisely the trade a placement decision must
weigh.

\begin{table}[t]
\caption{Linear fits $t(w)=t_0+s\cdot w$ of per-call time (ns) vs.\ callee work,
eight sizes. $t_0$ is the per-call boundary/dispatch intercept, $s$ the
per-iteration execution slope. Absolute $t_0$ is an extrapolation (small
negative values are low-work loop curvature); the load-bearing quantity is the
intercept \emph{gap} between twinned variants (see text). For \code{iop\_call} and \code{iop\_cb},
\code{free} executes in \code{sv} and \code{bridge} in \code{na}; the
$\approx$$4.4$ vs.\ $\approx$$184$\,ns/it slopes read off the executing
codespace. The lone low $R^2$ ($.912$, \code{iop\_call} bridge) is the
native path measured near the timer noise floor: at $\approx$$4.6$\,ns/it the
slope is real but low-SNR, and it is corroborated by three independent
\code{na}-resident series ($4.63/4.42/4.44$), so no single fit is load-bearing.}
\label{tab:fits}
\centering
\footnotesize
\setlength{\tabcolsep}{4pt}
\begin{tabular}{@{}llrrr@{}}
\toprule
Cell & Variant & $t_0$ (ns) & $s$ (ns/it) & $R^2$ \\
\midrule
\code{iop\_call} & free    & $-312$ & $86.69$  & $.9997$ \\
                 & bridge  & $461$  & $4.63$   & $.912$  \\
\midrule
\code{iop\_cb}   & free    & $-901$ & $182.18$ & $1.000$ \\
                 & bridge  & $4471$ & $183.53$ & $.9998$ \\
\midrule
\code{iop\_symmetric} & sv\_local  & $-387$  & $183.82$ & $1.000$ \\
                      & sv\_to\_na & $3184$  & $4.42$   & $.9998$ \\
                      & na\_local  & $123$   & $4.44$   & $1.000$ \\
                      & na\_to\_sv & $-1408$ & $188.49$ & $.9999$ \\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}[t]
\centering
\begin{tikzpicture}
\begin{semilogxaxis}[
  width=\columnwidth, height=4.4cm,
  log basis x=2,
  xlabel={callee work $w$ (LCG iterations/call)},
  ylabel={\code{iop\_cb} ratio},
  xtick={25,100,400,1600}, xticklabels={25,100,400,1600},
  ymin=0.95, ymax=1.8,
  ytick={1.0,1.2,1.4,1.6},
  grid=both, grid style={gray!20},
  tick label style={font=\scriptsize}, label style={font=\footnotesize},
  every axis plot/.append style={thick},
]
\addplot[mark=*,mark size=1.3pt,blue!70!black] coordinates {
  (25,1.732)(50,1.396)(100,1.286)(200,1.121)(400,1.102)(800,1.052)(1600,1.050)(3200,1.010)
};
\addplot[dashed,gray] coordinates {(25,1.0)(3200,1.0)};
\node[anchor=north east,font=\scriptsize] at (axis cs:3200,1.0) {no-cost (1.0)};
\end{semilogxaxis}
\end{tikzpicture}
\caption{The \code{iop\_cb} bridge/reference ratio decays toward $1.0$ as callee
work grows, because the callback crossing is a \emph{fixed} $\approx$$5.4$\,\textmu s
per-call cost (Table~\ref{tab:fits}), not a proportional overhead. The
$1.17\times$ headline is this constant seen at one small size.}
\label{fig:decay}
\end{figure}

\begin{figure}[t]
\centering
\begin{tikzpicture}
\begin{loglogaxis}[
  width=\columnwidth, height=5cm,
  log basis x=2,
  xlabel={callee work $w$ (iterations/call)},
  ylabel={per-call time (ns)},
  xtick={25,100,400,1600}, xticklabels={25,100,400,1600},
  legend pos=north west, legend style={font=\tiny, fill opacity=0.8},
  legend cell align=left,
  tick label style={font=\scriptsize}, label style={font=\footnotesize},
  every axis plot/.append style={thick, mark size=1.1pt},
]
\addplot[mark=*,red!70!black] coordinates {
  (25,4714)(50,9123)(100,17638)(200,35522)(400,71527)(800,148424)(1600,294560)(3200,587219)};
\addplot[mark=triangle*,orange!85!black] coordinates {
  (25,5206)(50,9632)(100,18564)(200,35180)(400,74398)(800,147026)(1600,295930)(3200,604403)};
\addplot[mark=square*,blue!70!black] coordinates {
  (25,3405)(50,3426)(100,3519)(200,4034)(400,5007)(800,6666)(1600,10273)(3200,17348)};
\addplot[mark=diamond*,teal!70!black] coordinates {
  (25,223)(50,347)(100,594)(200,1014)(400,1931)(800,3633)(1600,7175)(3200,14344)};
\legend{sv\_local,na\_to\_sv,sv\_to\_na,na\_local}
\end{loglogaxis}
\end{tikzpicture}
\caption{\code{iop\_symmetric}'s four placements. The two \code{sv}-executing
variants (\code{sv\_local}, \code{na\_to\_sv}) share the steep $\approx$$184$\,ns/it
slope; the two \code{na}-executing variants (\code{sv\_to\_na}, \code{na\_local})
share the shallow $\approx$$4.4$\,ns/it slope. \code{sv\_to\_na} sits a constant
$\approx$$3.2$\,\textmu s \emph{above} \code{na\_local}---that vertical offset is
the crossing intercept. Slope is placement; the offset is the boundary.}
\label{fig:placement}
\end{figure}

\subsection{Payload-cardinality scaling}
\label{sec:payload}

The family-1 sweep varies \emph{callee work} at fixed payload; the complementary
axis---\emph{payload size} at fixed work---is what turns a cross-runtime floor
into a per-element serialization curve. \code{xop\_feed\_payload} sweeps it: a
loopback \code{sv$\rightarrow$sv} RPC (\code{rpc}) and its in-process twin
(\code{direct}) each return a \code{list[$N$]} of \code{int} from the same
provider, with $N$ swept across sixteen sizes from $1$ to $100{,}000$ elements at a
fixed $20$ calls. Both sides run the identical \code{feed\_batch} LCG generator,
so at every $N$ the returned lists are byte-identical: the oracle passed at all
sixteen sizes (Table~\ref{tab:payload}), extending the differential-identity
witness across five orders of magnitude of payload and confirming the wire
round-trip preserves the list exactly.

Both paths fit $t(N)\approx t_0 + s\cdot N$ per call, but with opposite dominant
terms. The \code{direct} path is a near-perfect line of slope $s=298$\,ns/element
(95\% CI $296$--$299$, $R^2>0.9999$ over the whole range): building an
$N$-element list in-process costs essentially just per-element construction, with
no fixed floor to speak of. The \code{rpc} path is the mirror image---a
$\approx$$19$\,ms fixed per-call cost dominates at small $N$, and the per-element
serialization slope is buried beneath it until the payload grows large. Over the
full range the \code{rpc} fit is $s=934$\,ns/element (95\% CI $912$--$999$) on a
$\approx$$19$\,ms floor ($R^2=0.995$); the earlier sweep, which stopped at
$N{=}20{,}000$---essentially at break-even (below)---could only extrapolate this
slope, but the extended range now measures it directly, and cleanly: restricting
to the post-crossing band ($N\ge30{,}000$) gives $s=899$\,ns/element
(95\% CI $865$--$932$) on a $\approx$$22$\,ms floor at $R^2=0.9995$. That is the
result: \emph{crossing the RPC boundary costs $\approx$$3.1\times$ per element
what the in-process list costs} ($934$ vs.\ $298$\,ns/element)---the
JSON-encode/HTTP/decode marshalling tax laid bare as a slope, exactly the quantity
the family-1 method isolates, now measured on payload rather than work. The
consequence is visible in the ratio column of Table~\ref{tab:payload}: it
collapses from $27{,}000\times$ at $N{=}1$ to $3.7\times$ at $N{=}100{,}000$---the
inverse of the \code{iop\_cb} decay. There a fixed boundary was amortized by
growing \emph{work}; here a fixed RPC floor is amortized by growing \emph{payload},
while the direct twin it is measured against grows linearly.

Crucially, the crossing is now \emph{observed}, not inferred, and reported with
bootstrap confidence intervals rather than bare spread (percentile bootstrap,
$B{=}10^4$: paired round resampling for ratios, a pairs bootstrap over sweep points
for fit coefficients). The break-even payload at which the serialization term
equals the fixed cost is $N\approx21{,}000$ elements---almost exactly where the
earlier sweep stopped, which is why it could not see past the floor---and the
extended sweep measures $N$ nearly five times beyond it: at $N{=}100{,}000$ the
slope term ($\approx$$93$\,ms) is $\approx$$4.8\times$ the $\approx$$19$\,ms fixed
cost, so the crossing is firmly \emph{slope-dominated}. This inverts the earlier
reading---bounded at $20{,}000$, the crossing looked fixed-cost-dominated
throughout and the slope was a second-order extrapolation---and replaces it: the
loopback RPC is a fixed-cost floor at small payloads that hands off to a
serialization curve at large ones, with the handoff directly witnessed. The
confidence intervals put numbers on what was previously only
``order-of-magnitude'': the full-range slope is $934$\,ns/element (CI $912$--$999$,
a $\pm$$5\%$ interval), tightening to $865$--$932$ in the clean post-crossing
regime, while the per-point crossing medians stay wide at small $N$ (the $N{=}1$
\code{rpc}/\code{direct} ratio spans $23{,}600$--$31{,}700$). That small-$N$ width
is the fixed-cost floor's noise, not the slope's: the \code{rpc} medians are
non-monotone only for $N\le500$ and become strictly monotone from $N{=}1000$
onward---exactly where the slope emerges.

\begin{table}[t]
\caption{\code{xop\_feed\_payload}: per-call time returning a \code{list[$N$]}
over a loopback \code{sv$\rightarrow$sv} RPC vs.\ in-process, $N$ swept $1$ to
$100{,}000$ at $20$ calls, median of nine interleaved invocations. Every
row's \code{direct} and \code{rpc} canonical digests were byte-identical
($\checkmark$). \code{direct} fits $298$\,ns/element ($R^2{>}.9999$); \code{rpc}
fits a $\approx$$19$\,ms fixed cost plus a $934$\,ns/element serialization slope
(95\% CI $912$--$999$, $R^2{=}.995$), the slope directly observed to overtake the
fixed cost at $N\approx21{,}000$. Bootstrap CIs for every cell are in the released
\code{payload\_sweep\_results.json}.}
\label{tab:payload}
\centering
\small
\begin{tabular}{@{}rrrrc@{}}
\toprule
$N$ & \code{direct} (\textmu s) & \code{rpc} (ms) & rpc/direct & dig. \\
\midrule
1      & $0.69$   & $18.9$ & $27{,}480\times$ & $\checkmark$ \\
10     & $3.54$   & $19.7$ & $5{,}568\times$  & $\checkmark$ \\
50     & $15.5$   & $17.0$ & $1{,}095\times$  & $\checkmark$ \\
100    & $30.1$   & $20.2$ & $673\times$      & $\checkmark$ \\
250    & $74.1$   & $16.1$ & $217\times$      & $\checkmark$ \\
500    & $149$    & $17.7$ & $119\times$      & $\checkmark$ \\
1000   & $299$    & $18.3$ & $61\times$       & $\checkmark$ \\
2000   & $597$    & $23.1$ & $39\times$       & $\checkmark$ \\
5000   & $1{,}495$ & $28.8$ & $19\times$      & $\checkmark$ \\
10000  & $2{,}969$ & $29.3$ & $10\times$      & $\checkmark$ \\
20000  & $5{,}931$ & $40.2$ & $7\times$       & $\checkmark$ \\
30000  & $8{,}889$ & $48.0$ & $5\times$       & $\checkmark$ \\
40000  & $11{,}822$ & $58.4$ & $5\times$      & $\checkmark$ \\
50000  & $14{,}810$ & $66.6$ & $5\times$      & $\checkmark$ \\
75000  & $22{,}319$ & $88.7$ & $4\times$      & $\checkmark$ \\
100000 & $29{,}888$ & $111.6$ & $3.7\times$   & $\checkmark$ \\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}[t]
\centering
\begin{tikzpicture}
\begin{loglogaxis}[
  width=\columnwidth, height=5cm,
  log basis x=10,
  xlabel={returned payload $N$ (elements/call)},
  ylabel={per-call time (ns)},
  legend pos=south east, legend style={font=\tiny, fill opacity=0.8},
  legend cell align=left,
  tick label style={font=\scriptsize}, label style={font=\footnotesize},
  every axis plot/.append style={thick, mark size=1.1pt},
]
\addplot[mark=*,red!70!black] coordinates {
  (1,18933410)(10,19694205)(50,17019635)(100,20249846)(250,16068756)(500,17661349)(1000,18301284)(2000,23105189)(5000,28830003)(10000,29305711)(20000,40151207)(30000,47962922)(40000,58414764)(50000,66608500)(75000,88694946)(100000,111592625)};
\addplot[mark=square*,blue!70!black] coordinates {
  (1,689)(10,3537)(50,15544)(100,30078)(250,74120)(500,148883)(1000,299409)(2000,597080)(5000,1494784)(10000,2969337)(20000,5931120)(30000,8889142)(40000,11821696)(50000,14810078)(75000,22318661)(100000,29888315)};
\legend{rpc (crossing),direct (in-process)}
\end{loglogaxis}
\end{tikzpicture}
\caption{Payload-cardinality sweep, $N{=}1$ to $100{,}000$. \code{direct} is a
clean slope-$1$ line ($298$\,ns/element); \code{rpc} is a flat $\approx$$19$\,ms
floor that rises once the $934$\,ns/element serialization term climbs above it,
directly crossing break-even at $N\approx21{,}000$. The two curves converge as
$N$ grows---the crossing's fixed cost amortizes over a larger payload while the
in-process twin grows linearly---and by $N{=}100{,}000$ the crossing is
slope-dominated ($\approx$$4.8\times$ the fixed cost).}
\label{fig:payload}
\end{figure}

\section{Threats to Validity}
\label{sec:threats}

The differential-identity and structural-audit layers are the load-bearing
parts, and they are strong: byte-identical digests across every twin group and
matching manifest/wrapper facts are real correctness and compiler-regression
witnesses. The \emph{performance} layer is narrower than the kernel names
suggest, and we state its limits plainly.

\emph{Placement vs.\ boundary.} At a single size, \code{iop\_call} and
\code{iop\_symmetric} conflate \emph{which runtime runs the work} with crossing
cost. The size sweep (Section~\ref{sec:sweep}) separates the two---slope is
execution, intercept is crossing---but the separation is a linear extrapolation:
absolute intercepts are noisy (some slightly negative), so we quote only the
robust intercept \emph{gap} between twinned variants and the intercept of the
\code{sv\_to\_na} crossing, never a bare single-variant $t_0$.

\emph{FFI floors are not pure ABI.} \code{iop\_ffi\_scalar} includes two numeric
conversions per iteration; \code{iop\_ffi\_struct}'s $15$\,ns is an
\emph{average} across five struct shapes and must be split by shape before any
per-ABI-class claim.

\emph{Cross-runtime cells are floors, not workloads.} \code{xop\_feed} is a
\emph{sequential scalar} generated-client floor---no typed feed records and no
serialization-scaling measurement on that cell; it is not a HyperProtoBench-style
curve, and it measures a minimal custom \code{fetch} shim, not the shipped
\code{@jac/runtime}. The payload-scaling axis it lacks is supplied by the
separate \code{xop\_feed\_payload} sweep (Section~\ref{sec:payload}), which does
grow $N$ to $100{,}000$ elements over a loopback RPC; that cell in turn stays on a
\code{sv$\rightarrow$sv} loopback rather than a generated cross-language client,
so the two remain complementary floors rather than one production feed.
\code{xop\_wasm\_call} compares a Python-loop$\rightarrow$native-bridge against a
JavaScript-loop$\rightarrow$wasm-export---\emph{not} the same source built native
and wasm under equal host loops.

\emph{One machine; both axes now swept, payload gated by CIs.} The family-1 mixed-JIT
cells are swept across eight callee-work sizes and fit to a fixed-cost intercept
plus an execution slope (Section~\ref{sec:sweep}); the complementary
\emph{payload-bytes} axis, once future work, is now measured by
\code{xop\_feed\_payload} across sixteen cardinalities to $100{,}000$ elements
(Section~\ref{sec:payload}), separating the loopback RPC's $\approx$$19$\,ms fixed
cost from its $934$\,ns/element serialization slope. That cell is now gated by
bootstrap confidence intervals ($B{=}10^4$ percentile): the slope CI is
$912$--$999$\,ns/element and the crossing is directly observed to cross break-even
at $N\approx21{,}000$ and reach $\approx$$4.8\times$ slope-dominance by
$N{=}100{,}000$. One caveat remains: the per-point crossing medians are wide at
small $N$ (the $N{=}1$ ratio CI spans $23{,}600$--$31{,}700$), noise that lives in
the fixed-cost floor, not the slope. Family-2
scalar floors are median-of-seven with reported spread rather than a single
warm-up, but seven Node invocations still do not guarantee stable V8 tiering and
their crossing-side IQRs are wide. Outside the payload cell, results are one
machine and report spread without confidence intervals. None of these should be published as general Jac interop
overhead, realistic feeds, production microservices, or native-vs-wasm results.

We regard this candor as a feature. A benchmark suite for a language whose thesis
is measurable interop earns trust by naming exactly what each number licenses,
and by gating regressions on the facts it can defend (semantics and structure)
rather than the ones it cannot (absolute time on a shared runner).

\section{Design Implications}
\label{sec:implications}

The characterization is a means, not an end. Its purpose is to make the cost of
each boundary a quantity a compiler can \emph{act on}. Because a synechic
compiler owns both sides of every crossing, it already knows each call's
direction, static payload shape, and call-site frequency at compile time---the
exact inputs Table~\ref{tab:profiles} keys on. What the numbers here begin to
supply is the missing term: the per-crossing cost against which a placement or
marshalling decision pays off.

Three implications follow. The first two the sweep now quantifies; the third
remains a hypothesis. \emph{(i)~Placement is a slope; the crossing is an
intercept.} The sweep (Section~\ref{sec:sweep}) shows the two are literally
separable coefficients of one line: execution placement is a
$\approx$$42\times$ \emph{slope} difference ($4.4$ vs.\ $184$\,ns/iter) while the
crossing is a fixed \emph{intercept} ($\approx$$5.4$\,\textmu s for the
\code{iop\_cb} callback, $\approx$$3.2$\,\textmu s to enter \code{na}). A
compiler can therefore choose placement on callee slope alone once the intercept
is known small---and it can compute the exact break-even, which our fits make
concrete: entering \code{na} costs the $\approx$$3.2$\,\textmu s intercept and
saves $\approx$$179$\,ns/iter ($184$ minus $4.4$), so \emph{crossing into native
pays for itself above $\approx$$18$ iterations of callee work per call}. Below
that, stay in \code{sv}; above it, cross---a threshold computable at compile time
from the call's static work estimate. \emph{(ii)~The $\mu$s-vs-ns gap
sets a batching threshold.} A loopback RPC costs $10^{7}$--$10^{8}$\,ns per call
while an in-process dispatch costs $\sim$$10^{5}$; the break-even call count
above which batching or computation-migration wins is the ratio Table~%
\ref{tab:xruntime} brackets ($\sim$$10^2$). Applying the intercept/slope method
to \emph{payload} rather than work makes the two terms concrete
(Section~\ref{sec:payload}): the loopback RPC resolves into a $\approx$$19$\,ms
fixed cost plus a $934$\,ns/element serialization slope---$3.1\times$ the
in-process list's $298$\,ns/element---so a compiler that knows a call's static
payload cardinality can compute both when the crossing amortizes (large $N$) and
what each additional element on the wire will cost. \emph{(iii)~ABI class is a codegen choice,
not a constant.} The
$2$\,ns scalar vs.\ $15$\,ns struct-by-value gap is exactly the register-vs-%
\code{byval} distinction a compiler resolves per call signature. The decision
tree these imply---direct FFI, copied FFI, batched FFI, RPC, or migrate the
computation into the target codespace---is the natural next system contribution;
this paper supplies its measurement substrate and leaves the selector itself as
future work.

\section{Related Work}
\label{sec:related}

The Computer Language Benchmarks Game~\cite{clbg} and PolyBench~\cite{polybench}
compare language \emph{implementations} on shared kernels; JacInteropBench instead
holds the language fixed and varies the \emph{boundary}, so its unit is one Jac
program crossing one marshalled seam, not two languages solving one problem.
gRPC's benchmark methodology~\cite{grpc} separates fixed per-call cost from
per-byte work---the intercept/slope decomposition we implement over callee work
for the family-1 cells (Section~\ref{sec:sweep}) and over payload cardinality for
\code{xop\_feed\_payload} (Section~\ref{sec:payload}), which resolves a loopback
RPC into a $\approx$$19$\,ms fixed cost and a $934$\,ns/element serialization
slope, observed to cross break-even at $N\approx21{,}000$. HyperProtoBench~\cite{hyperproto} characterizes serialization at scale;
our payload sweep is a minimal instance of that payload-growing axis, still on a
loopback \code{sv$\rightarrow$sv} seam rather than realistic proto messages. TechEmpower~\cite{techempower} defines realistic web-endpoint
workloads that a future \code{xop\_crud} would slice. The manually engineered interop these boundaries replace has its own
tooling---\code{ctypes} and CFFI, pybind11~\cite{pybind11} and
PyO3~\cite{pyo3} for Python$\leftrightarrow$native, FastAPI and generated
OpenAPI clients for \code{cl}$\rightarrow$\code{sv}. A differential comparison
against these---same computation, same data representation, Jac-generated
crossing vs.\ hand-written glue---is the axis that would turn our
\emph{characterization} into a \emph{verdict} on compiler-generated interop. We
deliberately do not run it here: the honest contribution of this paper is the
twin-anchored oracle and the intra-Jac cost profiles, and a cross-system
comparison demands its own controls (identical kernels, matched marshalling,
per-system warm-up) to avoid attributing implementation differences to the
mechanism. It is the headline of the follow-on paper, not a claim we smuggle
into this one. Within Jac, \code{ownbench}
covers the free single-runtime boundary and the memory-management dial; this suite
is its marshalled-boundary complement, sharing its output contract, oracle
discipline, and compiler-regression role. The GraalVM/Truffle line~\cite{graal} is the closest prior ``one runtime, many
languages'' interoperability work, but it measures a shared managed VM's
polyglot call layer; JacInteropBench instead measures a compiler that lowers
\emph{to distinct runtimes} (Python bytecode, JavaScript, native/wasm) and so
must characterize genuinely heterogeneous boundaries. Our \code{xop\_wasm\_call}
caveat is precisely the confound Jangda et al.~\cite{jangda} isolate at scale
(wasm vs.\ native under matched host loops), which is why we report ours as an
asymmetric floor rather than a codegen verdict. On statistics, we follow the
spirit of rigorous performance evaluation~\cite{georges,renaissance} by
reporting full spread per cell, while deliberately declining confidence
intervals at $n{=}5$/$n{=}7$ rather than computing ones the sample cannot
support---and gating regressions on semantics, not time. To our knowledge no
established suite targets ``Python $\leftrightarrow$ WebAssembly
$\leftrightarrow$ native in one language,'' because until synechic compilation
there was no single language spanning those runtimes to measure.

\section{Conclusion}
\label{sec:conclusion}

Synechic compilation turns interop cost from an emergent property of hand-written
glue into a measurable property of one compiler. JacInteropBench is the
instrument for that measurement: ten measured kernels across two harness
families plus two named floors, each marshalled cell anchored to an exact
reference twin, a differential-identity oracle and structural audits gated in CI,
and provenance-stamped timing reported only within an explicitly bounded scope.
On commodity x86\_64 hardware an eight-point work sweep resolves per-call time
into a fixed crossing intercept ($\approx$$5.4$\,\textmu s for the callback
boundary, $\approx$$3.2$\,\textmu s to enter \code{na}) and a placement slope
($\approx$$42\times$, \code{na} vs.\ \code{sv})---so the $1.17\times$ callback
headline is a constant that decays to $1.01\times$ as the callee does real work---%
and a sixteen-point payload sweep to $N{=}100{,}000$ resolves the loopback RPC into a
$\approx$$19$\,ms fixed cost plus a $934$\,ns/element serialization slope
(95\% bootstrap CI $912$--$999$),
$3.1\times$ its in-process twin and directly observed to overtake the fixed cost
at $N\approx21{,}000$. We add low-nanosecond \code{C}-ABI foreign-call
costs, placement-dependent run-to-run variance spanning $1.5$ orders of magnitude,
and median-of-seven cross-runtime floors of $10$--$213\times$. The suite's most
durable output is not any single ratio but the guarantee that every crossing stays
byte-identical to its in-language reference on every commit. Future work fits the
remaining shape---the payload method carried onto a typed cross-language
\code{xop\_feed} feed, plus concurrency and RPS on
\code{xop\_crud} and an arm64 C-ABI runner---each landing behind the same
twin-anchored, regression-gated contract.

\begin{thebibliography}{9}
\bibitem{jac} Jaseci Labs, ``The Jac Programming Language,''
\url{https://github.com/jaseci-labs/jaseci}, 2026.
\bibitem{clbg} I.\ Gouy, ``The Computer Language Benchmarks Game,''
\url{https://benchmarksgame-team.pages.debian.net/benchmarksgame/}.
\bibitem{polybench} L.-N.\ Pouchet, ``PolyBench/C: The Polyhedral Benchmark
Suite,'' \url{https://web.cse.ohio-state.edu/~pouchet.2/software/polybench/}.
\bibitem{grpc} gRPC Authors, ``gRPC Performance Benchmarking,''
\url{https://grpc.io/docs/guides/benchmarking/}.
\bibitem{hyperproto} A.\ Kolesnikov et al., ``HyperProtoBench: A Realistic
Protocol Buffer Serialization Benchmark,'' Google, 2020.
\bibitem{techempower} TechEmpower, ``Web Framework Benchmarks,''
\url{https://www.techempower.com/benchmarks/}.
\bibitem{pybind11} W.\ Jakob, J.\ Rhinelander, and D.\ Moldovan, ``pybind11---%
Seamless operability between C++11 and Python,''
\url{https://github.com/pybind/pybind11}, 2017.
\bibitem{pyo3} PyO3 Project, ``PyO3: Rust bindings for Python,''
\url{https://github.com/PyO3/pyo3}.
\bibitem{jangda} A.\ Jangda, B.\ Powers, E.\ D.\ Berger, and A.\ Guha, ``Not So
Fast: Analyzing the Performance of WebAssembly vs.\ Native Code,'' in
\emph{USENIX ATC}, 2019.
\bibitem{graal} M.\ Grimmer, C.\ Seaton, R.\ Schatz, T.\ W\"urthinger, and
H.\ M\"ossenb\"ock, ``High-performance Cross-language Interoperability in a
Multi-language Runtime,'' in \emph{DLS}, 2015.
\bibitem{georges} A.\ Georges, D.\ Buytaert, and L.\ Eeckhout,
``Statistically Rigorous Java Performance Evaluation,'' in \emph{OOPSLA}, 2007.
\bibitem{renaissance} A.\ Prokopec et al., ``Renaissance: Benchmarking Suite
for Parallel Applications on the JVM,'' in \emph{PLDI}, 2019.
\end{thebibliography}

\end{document}
